---
title: "Nutrient Concentrations"
author: Kylie Newcomer
format:
  html: 
    output-file: "paper.html"
warning: FALSE

---
# Background
This goal of this project was to reproduce Figure 3 from "Effects of hurricane disturbance on stream water concentrations and fluxes in eight tropical forest watersheds of the Luquillo Experimental Forest, Puerto Rico. Journal of Tropical Ecology" by Schaefer et al 2000. This study analyzes the nutrient concentrations from streams of Bisley, Puerto Rico during 1986-1994, and analyzes the impacts of Hurricane Hugo on the concentrations.

```{r, echo = FALSE}
rm(list=ls())
library(tidyverse)
library(here)
library(janitor)
library(forecast)
library(zoo)
library(lubridate)
library(slider)
library(ARTofR)
library(paletteer)

```

# Data
The data used was taken from four sites in Bisley, Puerto Rico and measured the concentrations of Calcium, Potassium, Magnesium, Ammonium-N, and Nitrate-N throughout time. Raw data was combined to create one data set with all target sites. 
```{r}
# Reading in cleaned data
conc_final <- readRDS(here("outputs", "conc_final.RDS"))
```


# Methods
Data was filtered and selected to only include information necessary to run the analysis and apply a 9-week rolling average to the concentrations. The main steps after data wrangling and cleaning were to arrange samples by date, group them by site and nutrient type to run averages within individual sites and nutrients. The slide_index_dbl function was used to calculate the rolling mean over a 9-week period. 
```{r}
# Calling in my function
source(here("R", "rolling-mean-function.R"))
```


```{r}
# Applying function to data
conc_mean <- rolling_mean(conc_final)
```

# Results
The produced figure shows increases in concentrations of nutrients with available data across all sites following Hurricane Hugo in 1989. While most nutrient concentrations remained high in the years following the hurricane, Nitrate-N started to show a decrease after 1990.

```{r, fig.cap= "This figure shows the nutrient levels for rivers in Brisley, Puerto Rico with 1989 Hurricane Hugo denoted by the dotted line. Plots are broken up by the nutrients measured from top to bottom: Calcium, Potassium, Magnesium, Ammonium-N, and Nitrate-N"}
#Creating hurricane date variable for vline
hurricane <- as.Date("1989-09-18")

conc_mean_fig <- ggplot(conc_mean, aes(sample_date, roll_mean)) +
  geom_line(aes(color = sample_id)) +
  geom_vline(xintercept = hurricane, linetype = "dotted") + #adding huricane to graph
  scale_color_paletteer_d("nationalparkcolors::Arches") +
  facet_wrap(~nutrient, ncol = 1, scales = "free_y", strip.position = "left",
             labeller = as_labeller(
               c(ca = "Ca (mg l^-1)", 
                 k = "K (mg l^-1)", 
                 mg = "Mg (mg l^-1)", 
                 no3_n = "NO3-N (ug l^-1)", 
                 nh4_n = "NH4-N (ug l^-1)"))) +
  labs(x = "Years",
       y = "",
       color = "Site") +
  theme_minimal() +
  theme(strip.background = element_blank(), strip.placement = "outside", strip.text.y = element_text(size = 7.5), axis.text.y = element_text(size = 6))

conc_mean_fig
  
```

```{r}
ggsave(here("figs", "conc_mean_fig.png"), height = 5, width = 7)
```

